---
title: "R Notebook"
output: html_notebook
---

---
output:
  pdf_document: default
  word_document: default
  html_document: default
---
This is my first R Project, i want to use the knowledge of Time Series Analysis to Evaluate
and forcast the rate of covid-19 cases in Nigeria by the end of december 2020 using the
Prophet Packages in R. 
My data was pulled from https://ourworldindata.org/coronavirus-source-data
The data on the coronavirus pandemic is updated daily. Last update: November 3, 2020 (10:30, London time).

Reading the data file from my document into Covid19 variable
```{r}
covid19<- read.csv(file.choose())
```
lets check out the data set
```{r}
head(covid19, 10)
```

I want to try and get the column names
```{r}
colnames(covid19)
```
I will get rid off the unwanted columns
```{r}
covid19 <- covid19[ -c(6:49) ]
head(covid19,10)
```
I want to filter the data to get only cases related to Nigeria
```{r}
covidnaija<- covid19[covid19$location=="Nigeria", ]
head(covidnaija, 16)
```
```{r}
tail(covidnaija, 10)
```

i want to further remove the column on continent and iso-code, its not needed in my analysis
```{r}
covidnaija<-covidnaija[-c(1:2)]
head(covidnaija)
```
I want to back up filtered data so far in case
```{r}
covid_backup<-covidnaija
```
I want further check for missing data and NA data fields
```{r}
covidnaija[is.na(covidnaija$total_cases), ]
```
i have 70 rows of missing data most probably early days when corona virus cases was not recorded
I want to filter out these missing columns for my analysis.
```{r}
covidnaija<-covidnaija[!is.na(covidnaija$total_cases), ]
head(covidnaija)

```
```{r}
tail(covidnaija)
```
i want to check for missing data again
```{r}
covidnaija[is.na(covidnaija), ]
```
Great! no more missing data and columns
I want to reformat the indexing in the colunmns
```{r}
rownames(covidnaija)<-NULL
head(covidnaija)
```
Data is nice and clean and read to be used for forcasting.
For the Time series analysis, i will make use of a couple of libraries such as prophet, ggplot2, libridate etc.

```{r}
library("timeSeries")
library("prophet")
library("ggplot2")
library("lubridate")
library("forecast")
```
we need to format the date column into date format in R
```{r}
covidnaija$date<-ymd(covidnaija$date)
str(covidnaija)
```
Visualizing the The trend of Covid-19 cases in Nigeria
```{r}
qplot(data= covidnaija, date, total_cases,
      main= "COVID-19 confirmed Cases in Nigeria for 2020")
```

Passing Date and Prediction Variable into ds and y
```{r}
ds<-covidnaija$date
y<-covidnaija$total_cases
```

Creating a new data frame for the prediction vairables
```{r}
Newcovid<-data.frame(ds, y)
```

Passing the new data frame to Prophet for forecasting
```{r}
forcast1<-prophet(Newcovid)
```
Seasonality has been disabled for the above code because the data is not large enough to accomodate a whole season

i want to now create a data set predicted for the next 60 days
```{r}
predictions1<- make_future_dataframe(forcast1, periods = 60)
next60days<-predict(forcast1, predictions1)
```

Lets view the new future predictions 
```{r}
tail(next60days, 20)
```
So the new predicted data have predicted corona cases till 2nd of January 2021.
From Estimated and predicted data, Nigeria's Confirmed Corona Cases by the end of the year would slightly about 70,000 cases.

lets visualize this
```{r}
plot(forcast1, next60days)

```


```{r}
prophet_plot_components(forcast1, next60days)
```
The above code shows there are usually higher cases of Covid-19 cases in Nigeria during weekends starting from friday to sundays with Wednesdays having the lowest days of confirmed cases.
The high spike in Sundays and Saturdays could be attributed to people having to gather more in public places like market, churches or mosques, recreational centers e.t.c.

```{r}
predicted_values<-next60days$yhat[1:239]
Actual_values<-forcast1$history$y
plot(Actual_values, predicted_values)
```


